Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2020-05-06 docs/00-manuscripts/ieee/ieee-manuscript.tex

	\acro{AGB}{above-ground biomass}
	\acro{ALE}{accumulated local effects}
	\acro{ALS}{airborn laser scanning}
	\acro{ANN}{artificial neural network}
	\acro{AUROC}{area under the receiver operating characteristics curve}
	\acro{BRT}{boosted regression trees}
	\acro{CART}{classification and regression trees}
	\acro{CNN}{convolutional neural networks}
	\acro{CV}{cross-validation}
	\acro{DAP}{digital aerial photogrammetry}
	\acro{ENM}{environmental niche modeling}
	\acro{FPR}{false positive rate}
	\acro{FFS}{forward feature selection}
	\acro{FS}{feature selection}
	\acro{GAM}{generalized additive model}
	\acro{GBM}{gradient boosting machine}
	\acro{GLM}{generalized linear model}
	\acro{ICGC}{Institut Cartografic I Geologic de Catalunya}
	\acro{IQR}{interquartile range}
	\acro{MARS}{multivariate adaptive regression splines}
	\acro{MBO}{model-based optimization}
	\acro{MEM}{maximum entropy model}
	\acro{ML}{machine learning}
	\acro{NDII}{normalized difference infrared index}
	\acro{NIR}{near-infrared}
	\acro{NRI}{normalized ratio index}
	\acro{OLS}{ordinary least squares}
	\acro{LiDAR}{light detection and ranging}
	\acro{LOWESS}{locally weighted scatter plot smoothing}
	\acro{PISR}{potential incoming solar radiation}
	\acro{PCA}{principal component analysis}
	\acro{PDP}{partial dependence plots}
	\acro{PLS}{partial least-squares}
	\acro{RBF}{radial basis function}
	\acro{RF}{random forest}
	\acro{RMSE}{root mean square error}
	\acro{RR}{ridge regression}
	\acro{RSS}{residual sum of squares}
	\acro{SAR}{synthetic aperture radar}
	\acro{SDM}{species distribution modeling}
	\acro{SMBO}{sequential-based model optimization}
	\acro{SVM}{support vector machine}
	\acro{TPR}{true positive rate}
	\acro{VI}{vegetation index}
	\acro{XGBoost}{extreme gradient boosting}
\end{acronym}
\renewcommand{\IEEEiedlistdecl}{\relax}% reset back

@@

% mention why FS is used / important
Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
Feature selection approaches can help to address this issue, reducing possible in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.

@@

\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}) within the hyperspectral data, a buffer of two meters around the centroid of each tree was used for extraction of the reflectance values.
These averaged reflectance values were used as the base to derive all feature sets.
A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
In total, \(\frac{121*122}{2} = 7471\) NRIs were calculated due to the first four bands of the sensor being corrupted.

\subsection{Feature selection}

@@

To reduce the dimensions of a dataset, conceptually different approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or approaches like \ac{PCA} \cite{das2001, jolliffe2016}.
Since the main focus of this study is on investigating the effect of filter methods on predictive performance, only this sub-group will be introduced in more detail in the following.

@@

% Filter methods
The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
Some filter methods are restricted towards specific types of variables (numeric or nominal).
Filters do only rank features, they do not choose which covariates to drop or keep \cite{drotar2015}.
The selection of features to keep for the model fitting step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
Essentially, the number of covariates in the model is treated as a hyperparameter of the model.
The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.

% Ensemble filter methods
Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' are able to enhance the predictive performance of the resulting models \cite{abeel2010, drotar2017a}.
This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
For this filter, the final order is the sum of all single filters.

% Ensuring a fair weighting in the ensemble
Filter methods can be grouped into classes: Correlation-based, entropy-based, linear and non-linear methods.
It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased towards these.
In this study this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
In case pairs of filters with a correlation above 0.9 were discovered, only one of these was included into the ensemble filter.
By this we ensured that the ensemble filter is not biased towards a certain group of methods yielding highly similar rankings.

\subsubsection{Description of used filter methods}

@@

@@

\begin{itemize}
	\item  Extreme Gradient Boosting (XGBoost)
	\item  Random Forest (RF)
	\item  Penalized Regression (both L1 (lasso) and L2 (ridge))
	\item  Support Vector Machine (SVM)
\end{itemize}

\noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
\ac{XGBoost} (commonly abbreviated as XGBoost) showed promising results in benchmark competitions in recent years.
Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model\cite{hastie2001}.
Common penalties are 'lasso' (L1) and 'ridge' (L2).
The latter does not allow the full removal of variables from the model (penalization to zero) while the former does.
Both penalties can also be combined.
The combined approach is called 'elastic net' but was not used in this work.

@@

\begin{itemize}
	\item The raw hyperspectral band information (HR): No feature engineering) %chktex 13
	\item Vegetation Indices (\ac{VIs}): Expert-based feature engineering)
	\item Normalized Ratio Indices (\ac{NRIs}): Data-driven feature engineering)
\end{itemize}

@@

\noindent Estimating feature importance for highly-correlated datasets is a complicated task.
The correlation between covariates makes it challenging to calculate an unbiased estimate for single features.
Methods like \ac{PDP} or permutation-based approaches do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
Permutation feature importance calculates the loss of predictive performance for each feature by permuting it in a random manner.
The more important the feature is for the model, the higher the mean loss in the measure, reflecting the importance of the specific feature for the current model.
Predicting and calculating the error measure multiple times per feature comes with some computational cost.
For increased efficiency, we estimated the feature importance only for datasets HR and VI using the best performing learner of the benchmark.

@@

In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset, with a focus on the former.
With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while trying to avoid the over-interpretation of results.
We used the best-performing algorithm from the HR task (i.e. SVM) to calculate the feature importance.
The ALE plots show a small relative change compared to their respective mean effects for all chosen features ($< \mid 0.001 \mid$) (\autoref{fig:fi-hr-ale}).
Most ALE curves show a high dynamic across the value range of each variable, within their respective effect range.

@@

% - RMSE around 27
% - MBO tuned penalized methods are substantially worse
\noindent Overall, the response variable 'tree defoliation' could be modeled with an \ac{RMSE} of 28 \% percentage points (p.p.).
SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
Ridge faced major issues in four tasks due to one observation that was predicted off the scale.
SVM showed the best performance of all learners with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 for learner SVM (without filter) but up to 54.26 p.p. when testing on Laukiz2 (\autoref{tab:svm-single-fold-perf}).

@@

% learner
% - SVM Car is the overall winner
% - Ridge-CV hit and miss (hit = NRI, miss = VI)
% - all MBO penalized methods have no difference across tasks
SVM combined with the 'Carscore' filter achieved the best overall performance (RMSE of 27.98 p.p.) (\autoref{tab:perf-top-10}).
%L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In four out of six tasks (all which VI variables and HR-NRI) the error was out of bounds (\autoref{tab:perf-worst-10}).
For NRI and HR RMSE was 31.16 and 35.45 p.p., respectively.
In all settings for which ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
The specific observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
This one outlier caused the error estimates of these folds and the average estimate across all folds to be off the scale.

@@

\subsubsection{Model differences}
An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
These cluster around a performance of 31 p.p while SVM is able to score about 3 p.p. better than all other methods.
Comparing these differences (both relatively and absolute) to other studies would not be fair since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).

Penalized methods showed promising performances, especially when taking runtime into account.
When removing features with a correlation of 1, lasso is able to score performances around 31 p.p. and shows somewhat surprisingly the best performance on the VI task.
The same general conclusion applies to ridge if one discards the problematic results on four tasks.
A careful inspection of the hyperparameter optimization procedure of these models did not reveal any implementation mistakes.
One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
We therefore left it in the model, but we acknowledge that these learners seemed particularly sensitive to the input data.

\subsubsection{Feature set differences}
One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
With respect to \autoref{fig:perf-result}, no overall trend was found for all models that related to specific feature sets.
The performance of RF and XGBoost on the VI feature set was about six percentage points lower than on others.
One reason could be the lack of coverage in the wavelength area between 810 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
In addition, for all learners but SVM a better performance was observed when NRI indices were included in the feature set (i.e. NRI-VI, HR-NRI, HR-NRI-VI).

\subsection{Performance vs.\ plot characteristics}
\label{subsec:perf-plot-char}
The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
In this specific case, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
This caused overfitting on the medium-high values, degrading predictive performance in other scenarios.
In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).

@@

% - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
% - Filters can also have a negative impact
% - Filters are interesting for reducing runtime and improve interpretability, not only for performance
The usefulness of filters in this study varied.
While some filters were able to enhance model performances (up to 5 p.p. for RF and XGBoost), some caused a worse performance than using no filter at all \autoref{fig:filter-effects-no-filter}.
Since these negative cases were not caused by a specific filter method, one should include multiple filters in a modeling design if filters are going to be used.
While filters can improve the performance of models, they might be of more interesting in other aspects than performance: Reducing variables can reduce computational efforts in high-dimensional scenarios and might simplify interpretability of models.
Here, filters are a lot cheaper than wrapper methods due to their ability of being treated as a hyperparameter during the optimization stage.

% - Ensemble FS shows no advantage to simple methods -> always one simple filter which is better. Ensemble filters are good across many different datasets
The models which used the Borda ensemble method in this study did not score better on average than models which used a single filter or no filter at all.
Ensemble methods have higher stability and robustness than single ones and have shown promising results in \cite{drotar2017a}.
Hence, their main advantage are stable performances across datasets with varying characteristics.
Certain single filter methods might yield better model performances for certain datasets but fail on others.
The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in almost all cases (besides XGBoost on task HR-NRI-VI) a single filter outperformed the ensemble filter.
However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
Therefore no general statement can yet be made if the use of ensemble filters leads to a substantial better performance than applying one of the various single filter methods.
More case studies applying ensemble filter methods are needed.
We believe that ensemble filters can be a promising option in a machine-learning feature-selection portfolio.

% - PCA shows similar performance but probably requires tuning of main components -> check runtime!
PCA used as a filter in this work did neither show the best nor worst result compared to other feature-selection approaches.
It was able to reduce model fitting times substantially due to its nature of reducing the number of predictors to a minimum.
However, it also removed the possibility to conduct a post-hoc analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
Since filter scores only need to be calculated once in a benchmark setting, the runtime advantage might in fact be negligible in practice.

@@

The decision to use a buffer of two meters for the extraction of an observation's reflectance value was a complex decision.
Due to the reported geometric offset of up to 1 m within the hyperspectral data, the possibility of assigning a value to an observation which would actually refer to a different observation was reasonably high.
By using a buffer of two meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
However, the difference between single pixel values might often be higher than the difference between a smoothed value composed out of a buffer.
In our view using a buffer of two meters was a good compromise between including information of too many surrounding trees and not accounting for the geometric offset at all.
Even though no results showing the influence of multiple buffer values on the extraction were provided, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.

@@

The magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
Such limitations and uncertainties apply to many environmental studies and cannot be completely avoided.

@@

This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and their respective opposites \cite{horler1983}.
It is also worth mentioning that four out of ten of the most important features of dataset HR lie between 920 nm and 1000 nm.
Looking at the spectral curves of the plots, one can observe some apparent reflectance differences in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.

@@

% ALE plots
The ALE plots are hard to interpret due to their semi-meaningful absolute values which differ for each band.
For example, a reflectance value of 100 has a very different meaning in B67 than in B124 because for the latter reflectance values of vegetation are naturally higher.
Hence, a reflectance value of 100 would be considered high for B67 but low for B124 with respect to a spectral curve of vegetation.
However, no baseline reflectance value for every feature exists since the absolute reflectance values depend on the sensor characteristics.
Hence, it is hard to make reasonable interpretations of harsh drops/jumps within curves of certain features, e.g. the one at value 240 for feature B115.
ALE plots can be a powerful tool for interpreting feature importances of correlated features but might be limited in interpretability if the absolute feature values are only semi-meaningful.

@@

Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR}\cite{meng2018, kalin2019}.
\cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
However, both did not use spatial \ac{CV} and \cite{kalin2019} no \ac{FS}.
\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
Study results indicated that spectral metrics were found to be most helpful for the model.
Incorporating such metrics (both spectral and structural) could be a possible enhancement for future works.

% write about hyperspectral remote sensing
The field of (hyperspectral) remote sensing has a strong focus on using random forest for modeling in recent years \cite{belgiu2016}.
In high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computational expensive.
To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
However, we could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
Potential reasons could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.
Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for upcoming modeling studies using hyperspectral remote sensing data.

% RS studies with more/other algorithms than RF	
Remote sensing studies comparing multiple models often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.

>>>>

	\acro{AGB}{above-ground biomass}
	\acro{ALE}{accumulated local effects}
	\acro{ALS}{airborn laser scanning}
	\acro{ANN}{artificial neural network}
	\acro{AUROC}{area under the receiver operating characteristics curve}
	\acro{BRT}{boosted regression trees}
	\acro{CART}{classification and regression trees}
	\acro{CNN}{convolutional neural networks}
	\acro{CV}{cross-validation}
	\acro{DAP}{digital aerial photogrammetry}
	\acro{ENM}{environmental niche modeling}
	\acro{FPR}{false positive rate}
	\acro{FFS}{forward feature selection}
	\acro{FS}{feature selection}
	\acro{GAM}{generalized additive model}
	\acro{GBM}{gradient boosting machine}
	\acro{GLM}{generalized linear model}
	\acro{ICGC}{Institut Cartografic I Geologic de Catalunya}
	\acro{IQR}{interquartile range}
	\acro{MARS}{multivariate adaptive regression splines}
	\acro{MBO}{model-based optimization}
	\acro{MEM}{maximum entropy model}
	\acro{ML}{machine learning}
	\acro{NDII}{normalized difference infrared index}
	\acro{NIR}{near-infrared}
	\acro{NRI}{normalized ratio index}
	\acro{OLS}{ordinary least squares}
	\acro{LiDAR}{light detection and ranging}
	\acro{LOWESS}{locally weighted scatter plot smoothing}
	\acro{PISR}{potential incoming solar radiation}
	\acro{PCA}{principal component analysis}
	\acro{PDP}{partial dependence plots}
	\acro{PLS}{partial least-squares}
	\acro{RBF}{radial basis function}
	\acro{RF}{random forest}
	\acro{RMSE}{root mean square error}
	\acro{RR}{ridge regression}
	\acro{RSS}{residual sum of squares}
	\acro{SAR}{synthetic aperture radar}
	\acro{SDM}{species distribution modeling}
	\acro{SMBO}{sequential-based model optimization}
	\acro{SVM}{support vector machine}
	\acro{TPR}{true positive rate}
	\acro{VI}{vegetation index}
	\acro{XGBOOST}{extreme gradient boosting}
\end{acronym}
\renewcommand{\IEEEiedlistdecl}{\relax}% reset back

@@

Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
At the same time, it is necessary to account for the presence of spatial autocorrelation in the data.
The dataset used in this study comes with a spatial grouping at the plot level which needs to be accounted for \cite{sperrorest, pena2015}.

@@

\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}) within the hyperspectral data, a buffer of two meters around the centroid of each tree was used for extraction of the reflectance values.
A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
In total, \(\frac{121*122}{2} = 7471\) NRIs were calculated due to the first four bands of the sensor being corrupted.

\subsection{Dimension reduction}

\noindent We want to stress the differences between 'dimension reduction' and 'high-dimensionality'.
The former refers to the general idea of reducing features from a dataset \cite{vandermaaten2007}.
In modeling this means finding the best subset of covariates which provide the most predictive power to the model or extracting the main components of the features using a \ac{PCA}.
'High-dimensionality' in contrast is a dataset attribute and applies when \(p > n\), where \(p\) is the number of covariates and \(n\) the number of observations \cite{hastie2001}.
There is no absolute value at which the term applies, both for \(n\) and \(p\).
Hence for this study the word ‘high-dimensionality‘ only applies to the experiments using the ‘NRI‘ feature set ($>$ 7000 features for 1808 observations).

@@

\noindent In the following sections a brief overview of feature selection sub-categories is given.
Due to the focus of this study on the use of filter methods, which are a sub-group of feature selection methods, other approaches were grouped into a single section.

@@

% Filter methods
\noindent The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
Some filter methods are restricted towards specific types of variables (numeric or nominal).
Filters do only rank features, they do not choose which covariates to drop or keep \cite{drotar2015}.
The selection of features to keep for the model fitting step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
Essentially, the number of covariates in the model is treated as a hyperparameter of the model.
The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.

\paragraph{Ensemble filter methods}

% Ensemble filter methods
Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' are able to enhance the predictive performance of the resulting models \cite{abeel2010, drotar2017a}.
This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015}.
In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
For this filter, the final order is the sum of all single filters.

\paragraph{Ensuring a fair weighting in the ensemble}

Filter methods can be grouped into classes: Correlation-based, entropy-based, linear and non-linear methods.
It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased towards these.
In this study this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
In case pairs of filters with a correlation above 0.9 were discovered, only one of these was included into the ensemble filter.
By this we ensured that the ensemble filter is not biased towards a certain group of methods yielding highly similar rankings.

\paragraph{Description of used filter methods}

@@

\subsubsection{Wrapper methods and PCA}

\noindent Other approaches to assess feature importance are so called 'wrapper methods' and the \ac{PCA}\cite{das2001, jolliffe2016}.

% Wrapper approach
\noindent Wrappers \cite{chandrashekar2014, kohavi1997} apply search algorithms such as 'random search' or 'generic simulated annealing'.
First a (random) subset of features is chosen based on the selected algorithm.
In comparison to filters, no ranking is done in this step.
Now the model is fitted on the data and the performance is evaluated.
This is done multiple times, depending on the defined stopping criteria set by the user (performance, runtime, evaluations).
A disadvantage of this approach is that hyperparameter tuning can only be applied after the feature selection optimization finished.
Hence, the 'wrapper approach' is an expensive optimization method because two stages need to be run in sequential order.
Due to their high computational cost, wrappers were not considered in this work.

% PCA
A method with a completely different approach compared to filters and wrappers is the \ac{PCA}\cite{pearson1901, jolliffe2016}.
Here, the main components of the feature space are extracted and combined.
The number of principal components needed to explain a large portion of the variance is dynamic and depends on the dataset.
In practice, this number often ranges between two and five.
By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the majority of information available in the data.
This enables cheap model fitting with balanced loss of predictor information.
A disadvantage is the reduced interpretability due to the main components being a squashed version of the original covariates.

@@

\begin{itemize}
	\item  Extreme Gradient Boosting (XGBOOST)
	\item  Random Forest (RF)
	\item  Penalized Regression (both L1 (lasso) and L2 (ridge))
	\item  Support Vector Machine (SVM)
\end{itemize}

\noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
\ac{XGBOOST} (commonly abbreviated as XGBOOST) showed promising results in benchmark competitions in recent years.
Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model\cite{hastie2001}.
Common penalties are 'lasso' (L1) and 'ridge' (L2).
The latter does not allow the full removal of variables from the model (penalization to zero) while the former does.
Both penalties can also be combined.
The combined approach is called 'elastic net' but was not used in this work.

@@

\begin{itemize}
	\item The raw hyperspectral band information (HR): No feature engineering) %chktex 13
	\item Vegetation Indices (\ac{VI}): Expert-based feature engineering)
	\item Normalized Ratio Indices (\ac{NRI}): Data-driven feature engineering)
\end{itemize}

@@

% FIXME Talk about the bias of permutation FI in highly correlated settings
\noindent Estimating feature importance for highly-correlated datasets is a complicated task.
The correlation between covariates makes it challenging to calculate an unbiased estimate for single features.
Methods like \ac{PDP} do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
% FIXME: Add ref
Due to the noisiness of the dataset, estimates from model-agnostic approaches such as permutation-based feature importance should also be taken with care.
This method calculates the loss of predictive performance for each feature by permuting it in a random manner.
The more important the feature is for the model, the higher the mean loss in the measure, reflecting the importance of the specific feature for the current model.
Predicting and calculating the error measure multiple times per feature comes with some computational cost.
For increased efficiency, we estimated the feature importance only for datasets HR and VI using the best performing learner of the benchmark.

@@

In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset.
With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while trying to avoid the over-interpretation of results.
We used the best-performing algorithm from the HR task (i.e. SVM) to calculate the feature importance.
ALE plots were not discussed in greater detail and are available as supplementary material.

@@

% - RMSE around 27
% - MBO tuned penalized methods are substantially worse
\noindent Overall, the response variable 'tree defoliation' could be modeled with an \ac{RMSE} of 28 \% percentage points (p.p.).
SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
Ridge and lasso faced major issues up to the point of not achieving an error below 50 for datasets HR-NRI-VI and NRI-VI.
SVM showed the best performance of all learners with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 for learner SVM (without filter) but up to 54.26 p.p. when testing on Laukiz2 (\autoref{tab:svm-single-fold-perf}).

@@

% learner
% - SVM Car is the overall winner
% - Ridge-CV hit and miss (hit = NRI, miss = VI)
% - all MBO penalized methods have no difference across tasks
SVM combined with the 'Carscore' filter achieved the best overall performance (RMSE of 27.98 p.p.) (\autoref{tab:perf-top-10}).
%L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In four out of six tasks (all which VI variables and HR-NRI) the error was out of bounds (\autoref{tab:perf-worst-10}).
For NRI and HR RMSE was 31.16 and 35.45 p.p., respectively.
In all settings for which ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
The specific observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
This one outlier caused the error estimates of these folds and the average estimate across all folds to be out of bounds.

@@

\subsubsection{Model differences}
An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).

FIXME: Add refs when SVM was clearly worse then other algs.
In other (environmental) studies in recent years, SVM showed most often less good performances and was not able to outperform competing algorithms such as RF.
The difference to the second best performing model of around three percentage points is also worth noting:
We are not aware of a study in which SVM ranked as the best model that clearly.
All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).

We were a bit surprised about the performances of the penalized methods lasso and ridge.
The former was not able to achieve an RMSE below 46.0 p.p. across any feature set while ridge came at least somewhat close to RF and XGBoost for some feature sets.
However, there were also three settings (VI, NRI-VI, HR-NRI-VI) for which the RMSE of the ridge model was unreasonably large.
Careful inspection of the hyperparameter optimization of these learners did not reveal any implementation mistakes. However, one observation turned out to be influential even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
We therefore left it in the model, but we acknowledge that these learners seemed particularly sensitive to the input data.

FIXME: Talk about L1 and L2 model performance

\subsubsection{Feature set differences}
One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
With respect to \autoref{fig:perf-result} one can say that overall the different feature sets did not result in substantial and consistent changes in performance.
RF and XGBoost showed a somewhat surprising decrease in performance of about six percentage points on task VI compared to all others.
We have no explanation for this specific result, especially because the best performing SVM model does not reflect this difference.
In addition, all learners showed a slightly better average performance on feature sets which included NRIs (e.g. NRI-VI, HR-NRI, HR-NRI-VI) than on those which did not (i.e. VI and HR only).

\subsection{Performance vs.\ plot characteristics}
\label{subsec:perf-plot-char}

The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
In this specific case, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
This caused overfitting on the medium-high values, degrading predictive performance in other scenarios.
In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).

@@

% - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
The usefulness of filters in this study was apparent across all feature sets, even for the smallest feature set, HR.
Even though the effect of feature selection methods varied across algorithms, small to substantial improvement in predictive performance were achieved in almost all cases.
No optimal algorithm/filter/dataset setting was observed in this study, leading to the conclusion that one needs to try out various combinations of such to find the best performing one.

% - Ensemble FS shows no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
We have no explanation why the Borda ensemble method used in this study did not score better than the average performance of other filter methods.
FIXME: Add ref when and why Borda filters are good.
Ensemble methods have been proven to outperform simple filters in complicated scenarios when the latter face problems due to their focus on certain feature characteristics.
We suggest that the datasets used in this study were not complex enough to bring up these problems for simple features.
Therefore it makes sense that, due to its averaging nature, the Borda filter scored average results across all instances.
However, this cannot be known upfront and we believe that an ensemble filter should be part of every filter portfolio when doing filter-based feature selection.

% - PCA shows similar performance but probably requires tuning of main components -> check runtime!
The PCA approach did neither show the best or worst result when being compared to other feature-selection approaches.
It is able to reduce model fitting times substantially due to its nature of reducing the number of predictors to a minimum.
However, it also removes the possibility to conduct a post-analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
Since filter scores only need to be calculated once in a benchmark setting, the runtime advantage might in fact be negligible in practice.

@@

\noindent The decision to use a buffer of 2 m for the extraction of an observation's reflectance value was a complex decision.
Due to the reported geometric offset of up to 1 m within the hyperspectral data, the possibility of assigning a value to an observation which would actually refer to a different observation was reasonably high.
By using a buffer of 2 meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
However, the difference between single pixel values might often be higher than the difference between a smoothed value composed out of a buffer.
In our view using a buffer of 2 m was a good compromise between including information of too many surrounding trees and not accounting for the geometric offset at all.
Even though no results showing the influence of multiple buffer values on the extraction were provided, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.

@@

The magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
Such limitations and uncertainties apply to most applied studies and cannot be completely avoided.

FIXME: Talk about L1 and L2 models -> there we have quite some differences!

@@

This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and their respective opposites \cite{horler1983}.
It is also worth mentioning that four out of ten of the most important features of dataset HR lie between 920 nm and 1000 nm.
Looking at the spectral curves of the plots, we can observe quite some dynamic in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.

@@

@@

Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR}\cite{meng2018, kalin2019}.
\cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
Both study designs are substantially different compared to the setup of this work.
In addition, no spatial \ac{CV} or \ac{FS} was used.
\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
Study results indicated that spectral metrics were found to be most helpful for the model.
Incorporating such metrics (both spectral and structural) could be a possible enhancement for future works.

% write about hyperspectral remote sensing
The field of (hyperspectral) remote sensing has a strong focus on using random forest for modeling in recent years \cite{belgiu2016}.
In high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computational expensive.
To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
We could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
Potential reasons could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.
Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for upcoming modeling studies using hyperspectral remote sensing data.

% RS studies with more/other algs than RF	
Remote sensing studies comparing multiple models often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.
